---
title: "Feature Engineering"
author: "by A.G."
date: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  html_notebook:
    fig_width: 10
    fig_height: 6
    code_folding: hide
    df_print: paged
    toc: true
    number_sections: true
    toc_float: true
    theme: cerulean 
    highlight: kate
---

```{r, include=F}
knitr::opts_chunk$set(fig.width=10, fig.height=5)
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(ggpubr)
```

This is a notebook for creating various features and their exploration

# Make the main table

Which will contain all features and will be suitable for EDA

It should contain the following variables:

 - repeat length, 
 - repeat counts, 
 - AR length, 
 - center spanning, 
 - within gene location (of a repeat), 
 - distance to oriC,
 - density of matches,
 - IS presence (?),
 - IS length (=activity)
 - number of plasmids,
 - number of plasmids carrying RGs with repeats,
 - is an RG located on a plasmid,
 - "the black box" of AR
 
Its tidy structure should look like this:

Each strain contains:

  - 1 value for plasmid count, 
  - 1 value for a number plasmids carrying RGs
  - hundreds of genes; each gene (e.g. region) contains:
  
    - 1 distance to oriC, 
    - 1 density value, 
    - 'on a plasmid' label,
    - 1 value for IS count,
    - value for max IS length
    - a lot of repeat pairs; each repeat pair has:
    
        - 1 AR length, 
        - 1 AR black box vector
        - 1 repeat length, 
        - 1 center spanning label, 
        - 'inside the gene' label, 
 
```{r}

```
 
 