---
title: "RegulonDB overview"
author: "AG"
date: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float: true
    rmdformats:
      theme: readthedown
      highlight: github
---

# Intro

[RegulonDB](https://regulondb.ccg.unam.mx/) is the only functioning and maintained DB with operon data sets out there.

In **Downloads** section you can find a file `OperonSet.txt`

`wget https://regulondb.ccg.unam.mx/menu/download/datasets/files/OperonSet.txt`

There is a full version of the database available upon registration. I have registered but have no access to the full version.

Read the local version of `OperonSet.txt`

```{r}
library(readr)
# move OperonSet.txt to resources dir if the DB is useful
# the local copy is on GoogleDrive
col_names <- c("Operon_name", "First_gene-position_left", "Last_gene-position_right", "DNA_strand", "Number_of_genes",
               "Gene_name", "Evidence", "Confidence")
OperonSet <- read_delim("~/GoogleDrive/HeteroR/RegulonDB/OperonSet.txt", col_names = col_names, comment = "#",
                        delim = "\t", show_col_types = FALSE)
OperonSet
```

Type any R code in the chunk, for example:
```{r}
mycars <- within(mtcars, { cyl <- ordered(cyl) })
mycars
```

Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk.
Click the **Knit and Open Document** to build and preview an output.
