---
title: "Repeat Summary"
author: "AG"
date: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float: true
    rmdformats:
      theme: readthedown
      highlight: github
---

This notebook relies on results from *make_repeat_tables.py* that creates tables which contain
coordinates of repeat for each strain and each resistance gene

UPDATE: other results will also be here

# Make a table of all repeat tables

```{r, message = FALSE, warning = FALSE}
# a new way of getting repeat data frame across all strains
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)

strains <- dir("/home/andrei/Data/HeteroR/results/direct_repeats")
repeat_csv_paths <- sapply(strains, function(x){paste0("/home/andrei/Data/HeteroR/results/annotations/",x ,"/repeats/", x, "_repeats.csv")})

repeat_df <- bind_rows(lapply(repeat_csv_paths, function(x){read.csv(x)}))
# calculate AR length
repeat_df$AR_length <- repeat_df$start_2 - repeat_df$end_1

repeat_df
```

Number of rows: `r nrow(repeat_df)`

Number of strains: `r length(unique(repeat_df$strain))`

## Filter duplicates

There is no identical coordinate pairs (if you consider gene name as well)

```{r}
repeat_df_dedup <- repeat_df %>% distinct(record_id, start_1, end_1, start_2, end_2, .keep_all = TRUE)

```

Number of rows: `r nrow(repeat_df_dedup)`

Number of strains: `r length(unique(repeat_df_dedup$strain))`

## Sample 40 strains

To see fully interactive plots, go check `shiny_apps` or change the sample size and seed value.

```{r}
set.seed(99)

strains <- unique(repeat_df$strain)
strains_sample <- sample(strains, 40, replace = FALSE)

sample_df <- filter(repeat_df, strain %in% strains_sample)

sample_df
```



# Label repeats spanning their gene's center

The middle of the range is the middle of the RG.

Those repeats that are located on the same side of it, are not interesting for us because they don't affect amplifications rate

To get repeat coordinates I need bed files: regions_within.bed

The files are here:

`results/direct_repeats/DA*/regions/regions_within.bed`

```{r}
bed_df <- bind_rows(lapply(strains, function (x){
  filename <- paste0("/home/andrei/Data/HeteroR/results/direct_repeats/", x, "/regions/regions_within.bed")
  df <- read.delim(filename, header=FALSE)
  return(df)
}))

bed_df$gene_center <- round((bed_df$V3 - bed_df$V2 + 1)/2)
bed_df <- rename(bed_df, "record_id"=V4)
bed_df <- select(bed_df, record_id, gene_center)

bed_df
```

Number of gene names: `r length(unique(bed_df$record_id))`

Number of rows: `r nrow(bed_df)`

Number of unique gene names equals number of rows in bed_df, which means I can join bed and repeats tables using gene names only

## Joined repeat summary table and coordinates table

This table has been written to a file `/home/andrei/Data/HeteroR/results/repeats_summary_table.csv` and will be used in the Shiny app "repeats_statistic"

```{r}
repeat_df <- full_join(bed_df, repeat_df, by="record_id")
repeat_df$spans_center <- if_else(repeat_df$end_1 <= repeat_df$gene_center & repeat_df$start_2 >= repeat_df$gene_center, "yes", "no")

# write it to a file for that Shiny app
write.csv(repeat_df, "/home/andrei/Data/HeteroR/results/repeats_summary_table.csv")

repeat_df
```

# Analysis of repeat pairs

## Sample 40 strains

```{r}
sample_df <- filter(repeat_df, strain %in% strains_sample)

sample_df
```

## Repeat length

```{r, message=F, warning=F, fig.width=12}
box.rep.len <- plot_ly(data = sample_df,
                  x = ~strain,
                  y = ~length,
                  type = "box",
                  color=~spans_center) %>%
                layout(title="Repeat length",
                   boxmode="group",
                   yaxis=list(type="log"),
                   plot_bgcolor='#e5ecf6',
                   legend=list(title=list(text="Spanning gene's center")))

box.rep.len
```

## AR length distribution

```{r, message=F, warning=F, fig.width=12}
box.ar.len <- plot_ly(data = sample_df,
                      x = ~strain,
                      y = ~AR_length,
                      type = "box",
                      color=~spans_center) %>%
                layout(title="Length of amplifiable region", 
                       boxmode="group",
                       plot_bgcolor='#e5ecf6',
                       legend=list(title=list(text="Spanning gene's center")))

box.ar.len
```

## Repeat count per strain

```{r, message=F, warning=F}
repeat_counts_sample <- sample_df %>% group_by(strain, spans_center) %>% summarise("n_repeats"=n())
repeat_counts_sample
```

```{r, message=F, warning=F, fig.width=12}
bar.rep.counts <- plot_ly(data=repeat_counts_sample,
                          x = ~strain,
                          y = ~n_repeats,
                          type = 'bar',
                          color= ~spans_center) %>%
  layout(title="Number of repeats",
         boxmode="group",
         plot_bgcolor='#e5ecf6',
         legend=list(title=list(text="Spanning gene's center")))

bar.rep.counts
```


## Repeat count around each RG

### Make a table

```{r, message=FALSE, warning=F}

repeat_counts_gene_sample <- sample_df %>% group_by(strain, record_id, spans_center) %>% summarise("n_repeats"=n())

repeat_counts_gene_sample
```


### Make aplot

```{r, message=FALSE, warning=F, fig.width=12}
box.rep.counts.gene <- plot_ly(data = repeat_counts_gene_sample,
                      x = ~strain,
                      y = ~n_repeats,
                      type = "box",
                      color= ~spans_center) %>%
  layout(title="Repeat count around RGs",
         yaxis=list(type="log"),
         plot_bgcolor='#e5ecf6',
         boxmode="group",
         legend=list(title=list(text="Spanning gene's center")))

box.rep.counts.gene
```


## All together

```{r, message=F, warning=F, fig.height = 10, fig.width = 14}
annotations <- list(
  list(text="Repeat lengths", x=0.2, y=1.0, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE),
  list(text="Repeat counts", x=0.8, y=1.0, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE),
  list(text="Repeat counts around RGs", x=0.2, y=0.48, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE),
  list(text="AR lengths", x=0.8, y=0.48, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE))

subplot(box.rep.len, bar.rep.counts, box.rep.counts.gene, box.ar.len, nrows=2, shareX = T) %>% layout(showlegend=FALSE, annotations=annotations, title="")
```

# Statistics by main groups of AB

 - Penicillins - for example, phenoxymethylpenicillin, flucloxacillin and amoxicillin.
 - Cephalosporins - for example, cefaclor, cefadroxil and cefalexin.
 - Tetracyclines - for example, tetracycline, doxycycline and lymecycline.
 - Aminoglycosides - for example, gentamicin and tobramycin.
 - Macrolides - for example, erythromycin, azithromycin and clarithromycin.
 - Clindamycin.
 - Sulfonamides and trimethoprim - for example, co-trimoxazole.
 - Metronidazole and tinidazole.
 - Quinolones - for example, ciprofloxacin, levofloxacin and norfloxacin.
 - Nitrofurantoin - used for urinary infections.

Let's have a look at 40 strains that I have sampled before

## Number of repeats per AB group

Take AB classes from rgi tables? 
Or from known AB classes? 
Or group AB by mechanisms?

## RGI results table

Remove "Loose" hits

```{r}
# sampled strains n=40
#example_strains <- unique(repeat_counts_center2$strain)
#sample_strains <- sample(repeat_df$strain, 40)

# read in RG tables of these strains
rg_df <- bind_rows(lapply(strains_sample, function(x){
  df <- read.delim(paste0("/home/andrei/Data/HeteroR/results/resistance_genes/", x, "/rgi_table.txt"))
  df$strain <- x
  return(df)
}) )

ab_groups <- unique(rg_df$Drug.Class)

rg_df_notLoose <- rg_df %>% filter( Cut_Off != "Loose")
rg_df_notLoose
```

Number of rows in the full table `r nrow(rg_df)`, number of rows without "Loose" hits `r nrow(rg_df_notLoose)`

For *Strict* hits you may consider *Identity_percent* as another filtering criterion, hits with too low identities should be discarded.

```{r, message=F}
ggplot(rg_df %>% filter(Cut_Off == "Strict") %>% select(Best_Identities), aes(Best_Identities)) +
  geom_histogram(bins = 100, fill="steelblue") +
  ggtitle("Distribution of identity percentages among Strict hits")

```

## Number of entries in Drug.Class

Number of unique entries in Drug.Class: `r length(unique(rg_df_notLoose$Drug.Class))`

Unique entries in Drug.CLass

```{r}
unique(rg_df_notLoose$Drug.Class)
```

look at number of repeats around genes conferring resistance to each AB group

## Number of entries per AB group

```{r}
# "[M,m]etronidazole and [T,t]inidazole"
# "[S,s]ulfonamide[s] and [T,t]rimethoprim"
ab_groups <- c("penicillin", "cephalosporin", "tetracycline", "aminoglycoside", "macrolide", "clindamycin",
               "sulfonamide", "trimethoprim", "metronidazole", "tinidazole", "quinolone", "nitrofurantoin")

ab_counts_vec <- sapply(ab_groups, function (x){
  filter(rg_df_notLoose, grepl(x, Drug.Class, ignore.case = TRUE)) %>% nrow()
})

ab_counts_df <- data.frame("count" = ab_counts_vec, "AB.group" = names(ab_counts_vec))
rownames(ab_counts_df) <- NULL

ab_counts_df
```


```{r}
ab_found <- names(ab_counts_vec[ab_counts_vec != 0])
```

## Repeat counts per AB group

```{r}
# add AB labels to each gene in RGI table
rg_ab_found_df <- bind_rows(
  lapply(ab_found, function (x){
  df <- filter(rg_df_notLoose, grepl(x, Drug.Class, ignore.case = TRUE)) %>% select(ORF_ID, strain, Drug.Class)
  df$AB.group <- x
  return(df)
  })
)

# unify gene names in the repeats table
repeat_counts_gene_sample$record_id <- sub("_gene", "", repeat_counts_gene_sample$record_id)
rg_ab_found_df$record_id <- sapply(strsplit(rg_ab_found_df$ORF_ID, " "), function(x){x[1]})

# join repeat counts
count_per_ab <- left_join(repeat_counts_gene_sample, rg_ab_found_df, by="record_id") %>% filter(!is.na(Drug.Class))
count_per_ab
```

Boxplot

```{r, fig.width=16, fig.height = 8, message = FALSE, warning = FALSE}
box.rep.n.ab <- plot_ly(data = count_per_ab %>% filter(spans_center == "yes"),
                      x = ~strain.x,
                      y = ~n_repeats,
                      color = ~AB.group,
                      type = "box") %>%
  layout(title="Repeat counts around resistance genes",
         yaxis=list(type="sqrt", title="repeat count"),
         boxmode="group",
         xaxis=list(title=""),
         plot_bgcolor='#e5ecf6')

box.rep.n.ab
```

## AR length by AB group

On the same sample of strains

### Table

```{r}
# join repeat_df_center (contains AR lengths) and rg_ab_found (contains AB labels for each gene) by record_id
# prepare "record_id" column in repeat_df_center
repeat_df$record_id <- sub("_gene", "", repeat_df$record_id)
# join
repeat_df_center_ab_ar <- left_join(rg_ab_found_df, repeat_df %>% filter(spans_center == "yes"), by="record_id") %>% filter(!is.na(Drug.Class) & !is.na(spans_center))
repeat_df_center_ab_ar
```

### Boxplot

```{r, fig.width=16, fig.height = 8, message = FALSE, warning = FALSE}
box.ar.ab <- plot_ly(data = repeat_df_center_ab_ar,
                      x = ~strain.x,
                      y = ~AR_length,
                      color = ~AB.group,
                      type = "box") %>%
  layout(title="AR size variation per AB group",
         yaxis=list(type="sqrt", title="AR length"),
         boxmode="group",
         xaxis=list(title=""),
         plot_bgcolor='#e5ecf6')

box.ar.ab
```

## Repeat lengths variation by AB group

```{r, fig.width=16, fig.height = 8, message = FALSE, warning = FALSE}
box.rl.ab <- plot_ly(data = repeat_df_center_ab_ar,
                      x = ~strain.x,
                      y = ~length,
                      color = ~AB.group,
                      type = "box") %>%
  layout(title="Repeat length variation per AB group",
         yaxis=list(type="log", title="AR length"),
         boxmode="group",
         xaxis=list(title=""),
         plot_bgcolor='#e5ecf6')

box.rl.ab

```

## NB

This section might not be of a big value since we have Resistance/Susceptibility testing results in a separate data frame.


# Difference between resistant and susceptible strains

I have some (preliminary) resistance/susceptibility information for 400+ strains in `resources/resistance_testing.csv` (taken from Resilio storage).

```{r}
resistance_df <- read.csv("/home/andrei/Data/HeteroR/resources/resistance_testing.csv", na.strings = "")
resistance_df
```

Columns with *SIR* contain susceptibility test results, S - susceptible, R - resistant, I - intermediate.

---

CFX - Cefotaxime, Beta-lactam group

TZP - Pip-Taz

GM - Gentamicin, Aminoglycoside family

AMP - Ampicillin, Penicillin family, Aminopenicillin group

TBM - Tobramycin

TMP - Trimethoprim

NFT - Nitrofurantoin

MCN - Mecillinam

---

Let's leave only them in the data table

```{r}
resistance_df <- select(resistance_df, contains("SIR") | contains("strain"))
names(resistance_df) <- sub(".SIR", "", names(resistance_df))
resistance_df
```

```{r, message=FALSE, warning=FALSE}
resistance_df_td <- gather(resistance_df, key="AB", value="condition", 1:8)
resistance_df_td$AB <- as.factor(resistance_df_td$AB)
resistance_df_td$condition <- factor(resistance_df_td$condition, ordered=T, levels = c("S", "I", "R"))

# for plotly
condition_numeric <- sub("S", 0, resistance_df_td$condition)
condition_numeric <- sub("I", 1, condition_numeric)
condition_numeric <- sub("R", 2, condition_numeric)

resistance_df_td$condition_num <- as.integer(condition_numeric)

resistance_df_td
```

## Resistance heatmap

### Entire table

```{r, fig.width=14, fig.height=6, warning=F, nessage=F}
htmp_res_all <- ggplot(resistance_df_td, aes(strain, AB)) +
  geom_tile(aes(fill=condition)) +
  theme(axis.text.x = element_text(angle=45, size=4, vjust=0.5))


hm_res_all <- plot_ly(data = resistance_df_td,
                            x = ~ strain,
                            y = ~ AB,
                            z = ~ condition_num,
                            type = "heatmap") %>% 
    layout(title="Resistance testing results") %>% colorbar(title="Resistance level")

hm_res_all
```


### Sample subset

```{r}
res_df_sample <- filter(resistance_df_td, strain %in% strains_sample)
res_df_sample
```

```{r, fig.width=8, warning=F, nessage=F}
htmp_res_sample <- ggplot(res_df_sample, aes(strain, AB)) +
  geom_tile(aes(fill=condition)) +
  theme(axis.text.x = element_text(angle=45, vjust=0.5))


hm_res_sample <- plot_ly(data = res_df_sample,
                            x = ~ strain,
                            y = ~ AB,
                            z = ~ condition_num,
                            type = "heatmap") %>% 
    layout(title="Resistance testing results") %>% colorbar(title="Resistance level")

hm_res_sample
```

## Repeat number, length, AR length differences between R and S strains

I have genes conferring resistance to Gentamicin (GM) - *Aminoglycoside antibiotic* in the RGI results table.

Also, I have resistant and susceptible strains to GM in my collection - see the heatmaps above

Let's explore differences between R and S

I exclude I-strains since there's too few of them

...

### Prepare the data

```{r}
ab_res_repeat_df <- inner_join(resistance_df_td, repeat_df, by="strain") %>%
  filter(!is.na(condition)) %>%
  filter(condition != "I")

ab_res_repeat_df
```

### Repeat length

```{r, warning = F, message = F, fig.width = 14, fig.height = 6}

box.rl.res.y <- plot_ly(data = ab_res_repeat_df %>% filter(spans_center == "yes"),
                      x = ~AB,
                      y = ~length,
                      color = ~condition,
                      colors = c("#66C2A5", "#FC8D62"),  # taken from Set2
                      type = "box") %>%
                  layout(title="Lengths of repeats spanning region's center in resistant and susceptible strains",
                         yaxis=list(type="log", title="Repeat length"),
                         boxmode="group",
                         xaxis=list(title=""),
                         plot_bgcolor='#e5ecf6')

box.rl.res.n <- plot_ly(data = ab_res_repeat_df %>% filter(spans_center == "no"),
                      x = ~AB,
                      y = ~length,
                      color = ~condition,
                      colors = c("#66C2A5", "#FC8D62"),  # taken from Set2
                      type = "box") %>%
                  layout(title="Lengths of repeats not spanning region's center in resistant and susceptible strains",
                         yaxis=list(type="log", title="Repeat length"),
                         boxmode="group",
                         xaxis=list(title=""),
                         plot_bgcolor='#e5ecf6')

annotations <- list(
  list(text="Spanning center", x=0.2, y=1.0, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE),
  list(text="Not spanning center", x=0.85, y=1.0, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE))

subplot(box.rl.res.y, box.rl.res.n, nrows=1, shareX = T, shareY = T) %>% layout(showlegend=FALSE, annotations=annotations, title="Repeat lengths")

```

ggplot

```{r}
ggbox.rl.ab <- ggplot(ab_res_repeat_df, aes(AB , length)) +
  geom_boxplot(aes(fill=condition), varwidth=T, notch=T) +
  coord_trans(y="log") +
  facet_grid(spans_center ~ .) +
  scale_fill_brewer(palette = "Set2") +
  ggtitle("Repeat length")

ggbox.rl.ab
```


### AR length

```{r, fig.width = 8, fig.height = 5}
ggbox.arlen.res <- ggplot(ab_res_repeat_df, aes(AB , AR_length)) +
  geom_boxplot(aes(fill=condition), varwidth=T, notch=T) +
  facet_grid(spans_center ~ .) +
  scale_fill_brewer(palette = "Set2") +
  ggtitle("AR length")

ggbox.arlen.res
```

```{r, warning = F, message = F, fig.width = 8}

box.arlen.res.y <- plot_ly(data = ab_res_repeat_df %>% filter(spans_center == "yes"),
                      x = ~AB,
                      y = ~AR_length,
                      color = ~condition,
                      colors = c("#66C2A5", "#FC8D62"),  # taken from Set2
                      type = "box") %>%
                  layout(title="Amplifiable region length in resistant and susceptible strains",
                         yaxis=list(title="AR length"),
                         boxmode="group",
                         xaxis=list(title=""),
                         plot_bgcolor='#e5ecf6')

box.arlen.res.n <- plot_ly(data = ab_res_repeat_df %>% filter(spans_center == "no"),
                      x = ~AB,
                      y = ~AR_length,
                      color = ~condition,
                      colors = c("#66C2A5", "#FC8D62"),  # taken from Set2
                      type = "box") %>%
                  layout(title="Amplifiable region length in resistant and susceptible strains",
                         yaxis=list(title="AR length"),
                         boxmode="group",
                         xaxis=list(title=""),
                         plot_bgcolor='#e5ecf6')

annotations <- list(
  list(text="Spanning center", x=0.2, y=1.0, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE),
  list(text="Not spanning center", x=0.2, y=0.475, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE))

subplot(box.arlen.res.y, box.arlen.res.n, nrows=2, shareX = T) %>% layout(showlegend=FALSE, annotations=annotations, title="AR lengths")
```

### Repeat count

Prepare data

```{r, message=FALSE}
ab_res_rep_count <- ab_res_repeat_df %>%
  group_by(AB, condition, strain, spans_center) %>%
  summarize(n=n()) %>%
  filter(condition != "I")

ab_res_rep_count
```

Plot

```{r, warning = F, message = F, fig.width = 8}
ggbox.rc.res <- ggplot(ab_res_rep_count, aes(AB, n)) +
        geom_boxplot(aes(fill = condition)) +
        coord_trans(y="sqrt") +
        facet_grid(spans_center ~ .) +
        scale_fill_brewer(palette = "Set2")

ggbox.rc.res
```

```{r, mesaage=F, warning=F, fig.width=12, fig.height=6}
box.rc.res.y <- plot_ly(data = ab_res_rep_count %>% filter(spans_center == "yes"),
                      x = ~AB,
                      y = ~n,
                      color = ~condition,
                      colors = c("#66C2A5", "#FC8D62"),  # taken from Set2
                      type = "box") %>%
                  layout(title="Repeat numbers in resistant and susceptible strains",
                         yaxis=list(title="n"),
                         boxmode="group",
                         xaxis=list(title=""),
                         plot_bgcolor='#e5ecf6')

box.rc.res.n <- plot_ly(data = ab_res_rep_count %>% filter(spans_center == "no"),
                      x = ~AB,
                      y = ~n,
                      color = ~condition,
                      colors = c("#66C2A5", "#FC8D62"),  # taken from Set2
                      type = "box") %>%
                  layout(title="Repeat numbers in resistant and susceptible strains",
                         yaxis=list(title="n"),
                         boxmode="group",
                         xaxis=list(title=""),
                         plot_bgcolor='#e5ecf6')

annotations <- list(
  list(text="Spanning center", x=0.2, y=1.0, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE),
  list(text="Not spanning center", x=0.2, y=0.475, xref = "paper", yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE))

subplot(box.rc.res.y, box.rc.res.n, nrows=2, shareX = T) %>% layout(showlegend=FALSE, annotations=annotations, title="Repeat counts")
```


# Save workspace

```{r}
save.image(f="/home/andrei/Data/HeteroR/notebooks/repeats_summary.RData")
```
