---
title: "Modelling"
author: "by A.G."
date: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: hide
    fig_width: 10
    fig_height: 6
    theme: cerulean
    highlight: kate
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 5, message = F, warning = F)
library(dplyr)
library(caret)
library(car)
library(ggplot2)
library(GGally)
```

## Read data set

Data set `features_amp_strain` has been made previously in `feature_engineering.Rmd`

```{r}
feat_amp <- as_tibble(read.csv("tables/features_amp.csv"))
feat_amp <- rename(feat_amp, n.plasmids="n_plasmids") %>% relocate(n.plasmids, .before="n.rep.total")

# create dummy from original target
feat_amp$resistance_dummy <- ifelse(feat_amp$resistance=="R", 1, 0)
feat_amp <- relocate(feat_amp, resistance_dummy, .before="n.beta.lac")
# BL counts as an ordered factor
feat_amp$n.beta.lac.f <- factor(feat_amp$n.beta.lac, ordered = T, levels = c(1,2,3,4,5,6))
feat_amp <- relocate(feat_amp, n.beta.lac.f, .before="n.plasmids")
# resistance as factor
feat_amp$resistance <- as.factor(feat_amp$resistance)
feat_amp
```

Selected variables (based on plots in feature_engineering)

```{r, eval=FALSE}
feat_selected <- feat_amp %>% select(c(strain, resistance, n.beta.lac, n.beta.lac.f, n.plasmids, med.rep.len, med.tot.rep.len, max.rep.len, n.rep.total, med.AR.len.cen))

summary(feat_selected)
```

Response (resistance) classes are equally represented in the data set.

```{r, fig.width=10, fig.height=10}
ggpairs(feat_selected, 
        columns=c(3:10), 
        aes(color=resistance, alpha=0.2, dotsize=0.02), 
        upper = list(continuous = wrap("cor", size = 2.5)),
        diag=list(continuous ="barDiag"))+
  scale_color_brewer(palette = "Set1")+
  scale_fill_brewer(palette = "Set1")
```

# Model 0: Null model

```{r}
fit0 <- glm(formula = resistance_dummy ~ 1, data=feat_amp, family = binomial(link="logit"))

summary(fit0)
```


# Model 1: add BL-genes

Let's try to build a simple model and then run its diagnostics.

We assume that beta-lactamases is the main predictor of resistance (HR?). Let's check it out.

## BL genes number as discrete variable

```{r}
fit1a <- lm(formula = resistance_dummy ~ n.beta.lac, data=feat_amp, family = binomial(link="logit"))

summary(fit1a)

```

Credibility intervals:

```{r}
confint(fit1a)
```

Looks quite good: significant slope and intercept, F-statistic and CIs

## BL genes count as ordered factor

```{r}
# n.beta.lac+n.plasmids+med.rep.len+med.tot.rep.len+max.rep.len+n.rep.total+med.AR.len.cen
#prop.R <- (feat_amp %>% filter(resistance=="R") %>% nrow())/nrow(feat_amp)

fit1b <- glm(formula = resistance_dummy ~ n.beta.lac.f, data=feat_amp, family = binomial(link="logit"))

summary(fit1b)

```

None of the predictors is significant

But we know that chances a strain to be susceptible decrease sharply if number of BL genes in the strain is greater than 3

Let's code BL number in one binary variable 'greater than 3: yes or no'

### Recoding BL count to binary

```{r}
feat_amp$n.beta.lac.3 <- ifelse(feat_amp$n.beta.lac > 3, 1, 0)
feat_amp <- relocate(feat_amp, n.beta.lac.3, .before="n.plasmids")
```


```{r}
fit1c <- glm(formula = resistance_dummy ~ n.beta.lac.3, data=feat_amp, family = binomial(link="logit"))

summary(fit1c)
```

It looks much better!

### Models 0 and 1b comparison test

```{r}
anova(fit0, fit1c, test = "Chisq")
```

fit1c is better than fit0: residual deviance is lower, models are significantly different (p < 2.2e-16).

# Model 2: add AR size

median AR size of repeat pairs should work well according to boxplots.

## med.AR.len

```{r}
fit2a <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len, data = feat_amp, family = binomial(link = "logit"))

summary(fit2a)
```

```{r}
anova(fit0, fit1c, fit2a, test = "Chisq")
```

## med.AR.len.cen

for repeats spanning center it might be better but there are NAs. Replace them with 0

```{r}
feat_amp %>% select(resistance_dummy, n.beta.lac.f, med.AR.len.cen) %>% filter(is.na(med.AR.len.cen))
```

```{r}
feat_amp$med.AR.len.cen <- ifelse(is.na(feat_amp$med.AR.len.cen), 0, feat_amp$med.AR.len.cen)
```

```{r}
fit2b <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen, data = feat_amp, family = binomial(link = "logit"))

summary(fit2b)
```

```{r}
anova(fit0, fit1c, fit2b, test = "Chisq")
```

## med.AR.len.100

```{r}
fit2c <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.100, data = feat_amp, family = binomial(link = "logit"))

anova(fit2b, fit2c)
```

```{r}
fit2d <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen.100, data = feat_amp, family = binomial(link = "logit"))

anova(fit2b, fit2d)
```

filtered AR do not make the model better, plus they are correlated with med.AR.len.cen

```{r}
ggplot(feat_amp, aes(med.AR.len.cen, med.AR.len.cen.100))+
  geom_point(aes(color=resistance))

```


# Model 3: add maximum repeat length

## Not spanning center

```{r}
fit3a <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len, data = feat_amp, family = binomial(link = "logit"))

summary(fit3a)
```

```{r}
anova(fit2b, fit3a, test = "Chisq")
```

It gets better!

## Spanning center

```{r}
fit3b <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len.cen, data = feat_amp, family = binomial(link = "logit"))

summary(fit3b)
```

```{r}
anova(fit2b, fit3a, fit3b, test = "Chisq")
```

max.rep.len.cen doesn't decrease Res.Dev and the corresponding model is not signif. different.

# Model 4: add repeat counts

## Total

```{r}
fit4a <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len + n.rep.total, data = feat_amp, family = binomial(link = "logit"))

summary(fit4a)
```

```{r}
anova(fit3a, fit4a, test="Chisq")
```

No better

## Not spanning center

```{r}
feat_amp$n.rep.tot.cen <- ifelse(is.na(feat_amp$n.rep.tot.cen), 0, feat_amp$n.rep.tot.cen)

fit4b <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len + n.rep.tot.cen, data = feat_amp, family = binomial(link = "logit"))

summary(fit4b)
```

```{r}
anova(fit3a, fit4b, test="Chisq")
```
Not that much different...

Model 3a is the best so far

# Model 5: add number of plasmids

```{r}
fit5 <- glm(formula = resistance_dummy ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len + n.plasmids, data = feat_amp, family = binomial(link = "logit"))

summary(fit5)
```

Nope

# The winner is Model 3a

resistance ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len

```{r}
confint(fit3a)
```


## Validation

### TPR v FPR

```{r}
require(ROCR)

# predict prob
resistance.prob <- predict(object = fit3a, type = "response")
resistance.pred <- prediction(resistance.prob, feat_amp$resistance_dummy)

# calculate TPR and FPR
resistance.perf <- performance(resistance.pred ,"tpr","fpr")
plot(resistance.perf, colorize=T , xlab='false positive', ylab='true positive')
```

### Precision, Recall, F-score

```{r}
precision <- posPredValue(resistance.pred, feat_amp$resistance_dummy, positive="1")
recall <- sensitivity(predictions, y, positive="1")

F1 <- (2 * precision * recall) / (precision + recall)
```


### AUC

```{r}
auc  <- performance(resistance.pred, measure = "auc")
accuracy <- performance(resistance.pred, measure = "acc")

print(c("AUC", auc@y.values[[1]]))
```

### Accuracy

```{r}
plot(accuracy@x.values[[1]], accuracy@y.values[[1]], xlab=accuracy@x.name, ylab=accuracy@y.name)
```

Maximum accuracy `r max(accuracy@y.values[[1]])`

### Spec, sens, acc

```{r}
perf3  <- performance(resistance.pred, x.measure = "cutoff", measure = "spec")
perf4  <- performance(resistance.pred, x.measure = "cutoff", measure = "sens")
perf5  <- performance(resistance.pred, x.measure = "cutoff", measure = "acc")

plot(perf3, col = "red", lwd =2, xlab='prob', ylab='spec')
plot(add=T, perf4 , col = "steelblue", lwd =2)
plot(add=T, perf5, lwd =2)

legend(x = 0.6,y = 0.3, c("spec", "sens", "acc"), 
       lty = 1, col =c('red', 'steelblue', 'black'), bty = 'n', cex = 1, lwd = 2)
# determine x coordinate manually!
#abline(v= 0.165, lwd = 2, lty=6)
```

## Diagnostics

```{r,  fig.width=6, fig.height=4, eval=F}
qqPlot(fit3a, labels=feat_amp$strain, id.method="identify", simulate=T, main="Q-Q plot")

```



```{r}
durbinWatsonTest(fit3a)
```

```{r}
crPlots(fit3a)
```

```{r, eval=F}
ncvTest(fit3a)
```

# Decision Tree

```{r}
#library(party)

dt <- ctree(resistance ~ n.beta.lac.3 + med.AR.len.cen + max.rep.len, data = feat_amp)

```

```{r}
plot(dt)
```


------------------------------------------------------------------------

# Save the workspace

```{r, eval=TRUE}
save.image(f="/home/andrei/Data/HeteroR/notebooks/modelling.RData")
```
